Nail: A practical interface generator for data formats

### 2015-08-11 paper/thesis/design.tex
DNS CPP benchmark

\section{Design}
\label{s:design}

@@

\subsection{Overview}

@@

Nail grammars describe both the \emph{external format} and an
\emph{internal representation} of a protocol, as opposed to existing
grammar languages, which describe only the former and leave it up to
the programmer to construct an internal representation, typically in
the form of an abstract syntax tree.
Nail produces the following from a single, descriptive grammar:

\begin{CompactItemize}
\item \textit{type declarations} for the internal model,
\item the \textit{parser}, a function to parse a sequence of bytes into an
instance of the model, and
\item the \textit{generator}, a function to create a
sequence of bytes from an instance of the model.
\end{CompactItemize}

For example, Figure~\ref{fig:dns-full} shows an example grammar for DNS packets. Nail produces two
functions shown in Figure~\ref{fig:dns-full-api} that parse a DNS packet into the structures
partially shown in
Figure~\ref{fig:dns-full-struct} and vice versa.

@@

\paragraph{Semantic bijection.}
Parsing inputs and generating outputs form a bijection. However,
a bijection in the traditional sense often does not make sense for data
formats. Consider a grammar for a text language that tolerates white space, or a binary protocol
that tolerates arbitrarily long padding. Program semantics should be independent of the number of padding elements in the input, and Nail therefore
does not expose that information to the programmer. We call such discarded fields \emph{constants}. 
Similarly, programs should not necessarily preserve the layout of objects referred to by their offsets.

Therefore, Nail  maintains only a  \emph{semantic bijection} between the external format and the internal model.
That is, when Nail parses an input into an internal representation, and then
generates output from that representation, the two byte streams (input and
output) will have the same meaning (i.e., be interpreted equivalently by Nail).
However, the byte streams might not be identical. If the grammar
consists of a simple protocol without offset fields, constants, and the like, there is a
conventional bijection between  internal models and valid parser inputs.
%\footnote{Say, the physical layer of most communication
%  protocols is a possibly infinite sequence of symbols that are syntactically nil followed by a
%  pre-determined synchronization sequence and the actual contents of the transmission.}

\paragraph{Hide redundant information.}
Nail's internal model is designed to hide unneeded and redundant information from the application.
Nail introduces \emph{dependent fields}, which contain data that can be computed during generation
and need to be kept as additional state during parsing. Dependent fields are, for example, used to represent
lengths, offsets, and checksums. If dependent fields were exposed in the internal model, information
would be duplicated and inconsistent internal data structures could be produced when data is
modified. For example when using Nail to handle UDP packets, without dependent fields, programmers
might forget to update checksum fields when they modify the payload data.

\paragraph{Parser extensions.}
Real-world protocols contain complicated ways of encoding data. Fully representing these in an
intentionally limited model such as our parser language is impractical. Therefore, Nail introduces
\emph{transformations}, which allow arbitrary code by the programmer to interact with the parser and
generator. Nail parsers and generators interact with data through an abstract stream, which allows
reading and writing of bits as re-positioning. Transformations allow the programmer to write
functions in a general purpose language that consume streams and define new temporary streams, while
also reading or writing the values of dependent fields.

Initial versions of Nail's design included a special combinator for handling offset fields,
which consumed a dependent field and applied
a parser at the offset specified therein. However, it proved impossible to foresee all the ways in which a protocol could encode an offset;
for example, some protocols such as PDF and ZIP locate structures by scanning for a magic number starting at the
end of the file or a fixed offset. In nested grammars, offsets are also not necessarily computed from
the beginning of a file or packet.
Nail's transformations allow the programmer to write arbitrary functions that can
handle such structures and streams, which are a generic abstraction for input and
output data that allow the decoded data to be integrated with the rest of the
generated Nail parser.

@@

\subsection{Basics}

@@

\subsection{Dependent fields}
\label{s:dependent}
Data formats often contain values that are determined by other values or the layout of information,
such as checksums, duplicated information, or offset and  length fields.
We represent such values using \emph{dependent fields} and handle them transparently during
parsing and generation without exposing them to the internal model.

@@

Nail provides only one built-in combinator that uses dependent fields, \cc{n_of}, which acts like
the \cc{many} combinator, except it represents an exact number, specified in the dependent field, of
repetitions, as opposed as many repetitions as possible. For example, DNS labels, which
are encoded as a length followed by
a value are described in line 38 of the DNS grammar.
Other dependencies, such as offset fields or checksums, are not handled directly by combinators, but
through  transformations, as we describe next.

\subsection{Input streams and transformations}
\label{s:transforms}

>>>>

\chapter{Design}
\label{s:design}

@@

\section{Overview}

@@

Nail grammars describe both the \emph{external format} and an
\emph{internal representation} of a protocol.
Nail produces the following from a single, descriptive grammar:

\begin{itemize}
\item \textit{Type declarations} for the internal model, which
the application should use to represent data items in memory.

\item The \textit{parser}, a function that the application
should invoke to parse a sequence of bytes into an instance
of the above model.

\item The \textit{generator}, a function that the application
should invoke to create a sequence of bytes from an instance
of the model.
\end{itemize}

For example, Figure~\ref{fig:dns-full} shows a Nail grammar for DNS
packets.  For this grammar, Nail produces the type declarations
shown in Figure~\ref{fig:dns-full-struct}, and the parser and
generator functions shown in Figure~\ref{fig:dns-full-api}.

@@

\paragraph{Semantic bijection.}
Parsing inputs and generating outputs suggests a bijection
between external data and its internal representation. However,
a bijection in the traditional sense often does not make sense for data
formats. Consider a grammar for a text language that tolerates white space, or a binary protocol
that tolerates arbitrarily long padding. Program semantics should be independent of the number of padding elements in the input, and Nail therefore
does not expose that information to the programmer. We call such discarded fields \emph{constants}. 
Similarly, programs should not necessarily preserve the layout of objects referred to by their offsets.

Therefore, Nail  establishes only a  \emph{semantic bijection} between the external format and the internal model.
That is, when Nail parses an input into an internal representation, and then
generates output from that representation, the two byte streams (input and
output) will have the same meaning (i.e., be interpreted equivalently by Nail).
However, the byte streams might not be identical. If the grammar
consists of a simple protocol without offset fields, constants, and the like, there is a
conventional bijection between  internal models and valid parser inputs.
%\footnote{Say, the physical layer of most communication
%  protocols is a possibly infinite sequence of symbols that are syntactically nil followed by a
%  pre-determined synchronization sequence and the actual contents of the transmission.}

\paragraph{Hiding redundant information.}
Nail's internal model is designed to hide unneeded and redundant information from the application.
Nail introduces \emph{dependent fields}, which contain data that can be computed during generation
and need to be kept as additional state during parsing. Dependent fields are, for example, used to represent
lengths, offsets, and checksums. If dependent fields were exposed in the internal model, information
would be duplicated and inconsistent internal data structures could be produced when data is
modified. For example, when using Nail to handle UDP packets, without dependent fields, programmers
might forget to update checksum fields when they modify the payload data.

\paragraph{Parser extensions.}
Real-world protocols contain complicated ways of encoding data. Fully representing these in an
intentionally limited model such as our parser language is impractical. Therefore, Nail introduces
\emph{transformations}, which allow arbitrary code by the programmer to interact with the parser and
generator. Nail parsers and generators interact with data through an abstract stream, which allows
reading and writing of bits and re-positioning. Transformations allow the programmer to write
functions in a general-purpose language that consume streams and define new temporary streams, while
also reading or writing the values of dependent fields.

@@

Initial versions of Nail's design included a special combinator for handling offset fields,
which consumed a dependent field and applied
a parser at the offset specified therein. However, it proved impossible to foresee all the ways in which a protocol could encode an offset;
for example, some protocols such as PDF and ZIP locate structures by scanning for a magic number starting at the
end of the file or at a fixed offset. In nested grammars, offsets are also not necessarily computed from
the beginning of a file or packet.
Nail's transformations allow the programmer to write arbitrary functions that can
handle such structures and streams, which are a generic abstraction for input and
output data that allow the decoded data to be integrated with the rest of the
generated Nail parser.

\section{Basics}

@@

\section{Dependent fields}
\label{s:dependent}
Data formats often contain values that are determined by other values or the layout of information,
such as checksums, duplicated information, or offset and  length fields.
We represent such values using \emph{dependent fields} and handle them transparently during
parsing and generation without exposing them to the internal model.

@@

Nail provides only one built-in combinator that uses dependent fields, \cc{n_of}, which acts like
the \cc{many} combinator, except it represents an exact number, specified in the dependent field, of
repetitions, as opposed to as many repetitions as possible. For example, DNS labels, which
are encoded as a length followed by
a value, are described in line 38 of the DNS grammar.
Other dependencies, such as offset fields or checksums, are not handled directly by combinators, but
through  transformations, as we describe next.

\section{Input streams and transformations}
\label{s:transforms}

### 2015-08-11 paper/thesis/eval.tex
DNS CPP benchmark

\section{Evaluation}
\label{s:eval}

@@

\subsection{Data formats}
\label{s:eval-formats}

To answer the first question, we used Nail to implement grammars
for seven protocols with a range of challenging features.
Figure~\ref{fig:eval-protocols} summarizes these protocols, the lines
of code for their Nail grammars, and the challenging features that make
the protocols difficult to parse with state-of-the-art parser generators.
We find that despite the challenging aspects of these protocols, Nail is
able to capture the protocols, by relying on its novel features: dependent
fields, streams, and transforms.  In contrast, state-of-the-art parser
generators would be unable to fully handle 5 out of the 7 data formats.
In the rest of this subsection, we describe the DNS and Zip grammars in
more detail, focusing on how Nail's features enable us to support these
formats.

@@

\subsection{Programmer effort}
\label{s:eval-effort}

@@

\subsection{Security}

@@

\subsection{Performance}

@@

The results are shown in Figure~\ref{fig:perf-dns}, and demonstrate that
our Nail-based DNS server can achieve higher performance and lower latency
than BIND\@.  Although BIND is a more sophisticated DNS server,
and implements many features that are not present in our Nail-based DNS
server and that allow it to be used in more complicated configurations, 
we believe our results demonstrate that Nail's parsers are not
a barrier to achieving good performance.

\input{fig-perf-dns}

>>>>

\chapter{Evaluation}
\label{s:eval}

@@

\section{Data formats}
\label{s:eval-formats}

To answer the first question, we used Nail to implement grammars
for seven protocols with a range of challenging features.
Figure~\ref{fig:eval-protocols} summarizes these protocols, the lines
of code for their Nail grammars, and the challenging features that make
the protocols difficult to parse with state-of-the-art parser generators.
We find that despite the challenging aspects of these protocols, Nail is
able to capture the protocols, by relying on its novel features: dependent
fields, streams, and transforms.  In contrast, state-of-the-art parser
generators would be unable to fully handle 5 out of the 7 data formats.
In the rest of this section, we describe the DNS and Zip grammars in
more detail, focusing on how Nail's features enable us to support these
formats.

@@

\section{Programmer effort}
\label{s:eval-effort}

@@

\section{Security}

@@

\section{Performance}

@@

The results are shown in Figure~\ref{fig:perf-dns}, and demonstrate
that our Nail-based DNS server can achieve higher performance and
lower latency than BIND\@.  Although BIND is a more sophisticated DNS
server, and implements many features that are not present in our
Nail-based DNS server and that allow it to be used in more complicated
configurations, we believe our results demonstrate that Nail's parsers
are not a barrier to achieving good performance.  Previous papers on
Nail presented results using a Nail parser generated as C, whereas
these results were produced on the same configuration with the newer
C++ parser generator, which improves performance by using a more
efficient stream representation.
\input{fig-perf-dns}

### 2015-08-11 paper/thesis/relwk.tex
DNS CPP benchmark

\section{Related work}
\label{s:relwk}

\paragraph{Parsers.}

@@

We presented an earlier design of Nail at a
workshop~\cite{bangert:nail-langsec}.  At that stage, Nail had only
limited support for dependent fields, and did not support stream
transforms at all, which are crucial for supporting real-world formats
like DNS and ZIP\@.  The workshop paper also did not provide a detailed
design discussion or evaluation.

>>>>

\chapter{Related work}
\label{s:relwk}

This thesis is based on previously published papers about Nail, including \cite{bangert:nail-langsec}, \cite{bangert:nail-osdi14} and \cite{bangert:nail-login}.
We presented an earlier design of Nail at a
workshop~\cite{bangert:nail-langsec}.  At that stage, Nail had only
limited support for dependent fields, and did not support stream
transforms at all, which are crucial for supporting real-world formats
like DNS and ZIP\@.  The workshop paper also did not provide a detailed
design discussion or evaluation.
Since presenting the conference paper~\cite{bangert:nail-osdi14}, we implemented the C++ backend and explored novel uses of Nail for network protocols. 
\paragraph{Parsers.}

@@
